---
title: "Automatic Rename Pictures with R "
subtitle: "An experiment with exif and geolovation"
author: "Christophe Bontemps"
date: "2024-03-16"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: true
    highlight: tango
    number_sections: true
    number-depth: 1
    theme: lumen
  pdf_document: default
always_allow_html: true
---

```{r setup, include=FALSE}
library(knitr)
# Global options
knitr::opts_chunk$set(echo = FALSE)

# Other options used here 
opts_chunk$set(warning = FALSE, message = FALSE,
               fig.pos = "!H", fig.align = "center",
               autodep = TRUE, tidy = FALSE)
```


```{r libraries}
# To extract information from images 
library(exifr)

# data management 
library(dplyr)
library(janitor)
library(openxlsx)

# Tables
library(kableExtra)

# For maps and geocoding
library(tidygeocoder)
library(leaflet)
library(maps)

```



```{r exiftoolconfig}
#May be needed to find exiftool (in Windows directory or anywhere in the path)
# From https://exiftool.org/install.html

# library(exiftoolr)
# configure_exiftoolr()
```

# Parameters

## Basic parameters
There are basically 2 types of information

 * Folders for file copying
 * Type of files to be considered (*e.g.*  .jpeg, .png)

This is where I define source and destination, obviously. It is possible to use a list of folders for the source to avoid tedious repetitions (Todo)

```{r paths}
# We only rename  pictures and videos
mypattern  <- "*.jpg|*.JPG|*.PNG|*.MP4|*.MOV"

# Source directory
SourceDir <-  "c:/temp/Photos/FullTest/"
# Destination 
ToDir <- "c:/temp/Photos/Fulltest/Renamed/"

# Create destination folder if doesn't exit
if (!dir.exists(ToDir)) {
  dir.create(ToDir, recursive = TRUE)
}
```


```{r paths}
# List of  files that will be renamed
files <- list.files(path = SourceDir, pattern = mypattern,  full.names = TRUE)
# Count the number of files
NbFiles <- length(files)

```

We will transfer **`r NbFiles`** pictures from *`r SourceDir`* to several sub-directories in *`r ToDir`*.

**NB**: *One idea would be also to do batches of transfers*

## Special events
There may be special events that require a specific tag/folder (*e.g.*  festival, concert). This information should be stored in an external file. 

**Todo**

## Triggers for automatic detection of  events

```{r}

# Nb of pictures taken the same day  --> Signaling a touristic trip
NbPicturesDayMax <- 50

# Nb of pictures taken in the same City  --> Signaling region folder will be too large 
NbPicturesCityMax <- 50
```

## Parameters on the desired output

**Todo**
 - Define the minimum number of pictures per directory! 
 

# Extract data from pictures

## Extract exif info
Once exiftool is correctly installed^[See the information on how to install [https://exiftool.org/install.html](https://exiftool.org/install.html)], one can collect all the information embedded in our pictures. Here we only collect the files with extensions *`r mypattern`*.

We do this only if exif has not been done already for the same files

```{r exifextraction  }
# Construct unique exif Filename to save
exifFileName <- paste0(SourceDir,"exif",NbFiles,".Rdata")

# Listing of files
Myfiles <- list.files(path = SourceDir,
                      recursive=TRUE,
                      pattern = mypattern, full.names=TRUE)

# Check if the file exists
if (file.exists(exifFileName)) {
  # Load the file if it exists
  print("Loading existing Exif file ;-) ")
  load(exifFileName)
} else {
  print("Exif file does not exist: Extracting (can be long)")
  # This is long... 
  exifinfo <- read_exif(Myfiles, quiet = TRUE)
  
  print(paste("Saving exif file", exifFileName) )
  save(exifinfo, file = exifFileName)
}
```

We have now the complete information on **`r nrow(exifinfo)`** pictures, among which **`r nrow(subset(exifinfo, FileType =="JPEG"))`** are jpegs. 


```{r selecting variables }

# Using the data 
# https://www.r-bloggers.com/2016/11/extracting-exif-data-from-photos-using-r/
# and https://www.r-bloggers.com/2016/11/create-an-interactive-web-map-with-geotagged-photos/

exifdat <- exifinfo %>%
  select(
    SourceFile, FileName, FileType, FileSize, Duration,  FileModifyDate, DateTimeOriginal,
    GPSLongitude, GPSLatitude, GPSAltitude) %>%
  arrange(FileModifyDate)

```

## Refining the information

We only keep a few elements from the pictures' exif information. 

```{r}
head(exifdat)
```


We need a bit of **cleaning** there, in particular, we have date format and, as usual, this is needs to be extracted if one wants to use the year as a sorting variable. Also some coordinates (GPS) are missing and values should be imputed (as zeros here) for *maps* to work (see below).


```{r DateInfo}
library(lubridate)
library(maps)

# Adding date 
mydat <- exifdat %>%
  mutate(DateTimeOriginal = ifelse(is.na(DateTimeOriginal),
                                   FileModifyDate,
                                   DateTimeOriginal)) %>%
  mutate(Date = ymd_hms(DateTimeOriginal)
  ) %>%
  arrange(Date) %>%
 tidyr::separate(Date, c("Year", "rest"), extra = "merge", fill = "left")  %>%
 tidyr::separate(rest, c("Month", "rest2"), extra = "merge", fill = "left") %>%
 tidyr::separate(rest2, c("Day", "Time"),  extra = "merge", fill = "left") %>%
 mutate(
   Date =  paste0(Year,"-",Month, "-", Day),  
   NewName = paste0(Date, "-",  FileName))
```
## Identifying pictures that will be renamed

TODO: 
- Manage pictures with no issues
- Pictures with no good information --> decide what to do. 

# Getting Geolocalisation

```{r}
# Replace missing coordinates by (0,0)
mydat <- mydat %>%
  mutate(GPSLongitude =  ifelse(is.na(GPSLatitude), 0 ,GPSLongitude),
         GPSLatitude =  ifelse(is.na(GPSLatitude), 0 ,GPSLatitude)) 
```


We use package **tidygeocoder** (1.0.5) for getting this information 

```{r}
# Using library(tidygeocoder)
# https://jessecambon.github.io/tidygeocoder/articles/tidygeocoder.html

# Creating OSM unique name 
osmfileName <- paste0(SourceDir,"osmdata",NbFiles,".Rdata")

# Check if the file exists
if (file.exists(osmfileName)) {
  # Load the file if it exists
  print("Loading  existing osm file ;-) ")
  load(osmfileName)
} else {
  print("Reverse location using OSM data ")
  osmdata <- reverse_geo(lat = mydat$GPSLatitude, long = mydat$GPSLongitude,
                method = "osm",
                custom_query = list(`accept-language` = "en"),
                full_results = TRUE)
  print(paste("Print OSM data ",osmfileName ))
  save(osmdata, file = osmfileName)
}
  
# this operation is also very long...
```


```{r}
# Extracting OSM data and merging
revdata <- osmdata %>%
  select( "lat", "long",  "country", "country_code", "city", "town" ,"type", "place_id", "name",  "class", "type", "village", "municipality", "county",   "state", "neighbourhood", "suburb", "leisure", "address")%>%
  rename(osmcity = city, 
         osmcountry = country, 
         osmcode = country_code) %>%
  distinct(lat, long, .keep_all = TRUE)

mydatosm <- left_join(mydat, revdata, 
              by = join_by(GPSLatitude == lat, GPSLongitude == long ) )

```


###  Collecting country information 

We use the package *maps* to recover the information from GPS (latitude, longitude). We will use this information to organize the files into folder names with information on  years and  location. 

```{r CountryInfo}
# Using the maps library

# Adding Countries from  GPS 
mydatN <- mydatosm %>%
  mutate(CountryGPS = maps::map.where(database="world", GPSLongitude, GPSLatitude, na.rm=TRUE),
         DeptGPS = maps::map.where(database="france", GPSLongitude, GPSLatitude, na.rm=TRUE))

# Create CountryShort and Region for detailed information 
mydatN <- mydatN %>%
  tidyr::extract(CountryGPS, c("Countryshort", "RegionGPS"),
                 regex = "([[:alnum:]]+):([[:alnum:]]+)",
                 remove = FALSE)%>%
  mutate(CountryGPS = ifelse(!is.na(Countryshort) , Countryshort, CountryGPS), 
         RegionGPS = ifelse(!is.na(DeptGPS) , DeptGPS, RegionGPS) )%>%
 select( !c("Countryshort", "DeptGPS"))
```


```{r CountryInfoOld}
# Identifying specific places/events
### Identifying  P2000
# mydatN <- mydatN %>%
#   mutate(Country = ifelse(GPSAltitude >1700, "P2000", Country)) 
#  
# ####  Festa Design
# mydatN <- mydatN %>%
#   mutate(Region = ifelse(Date %in% c("2022-11-19", "2022-11-20"), "Design-Festa", Region)) 
  
```


```{r}
mydatN <- mydatN %>%
  mutate(Country = case_when( osmcode =="jp" ~ "Japan",
                              is.na(CountryGPS) ~ osmcountry,
                             .default =  CountryGPS),
         Region =  case_when((GPSAltitude >1700) & (CountryGPS =="France") ~ "P2000",
                             grepl("balma", address, ignore.case = TRUE)~ "Balma", 
                             Date %in% c("2022-11-19", "2022-11-20") ~ "Design-Festa", 
                             .default =  RegionGPS)
         )
  
                              
```

## Test on the outcome

TODO: Managing issues with this 

* Bad names or Japanese city with Japanese characters
* Not found location


### Visualizing Missing

```{r}
library(naniar)
vis_miss(mydatN, sort_miss =  TRUE )
```


>  We decide to use ***osmcity***, ***town*** and ***municipality*** to create information about the location ant the city level. . Seems to work wuite well! 
 
```{r}
# Define Newcity based on missing 
mydatN <- mydatN %>%
  mutate(Newcity = ifelse(is.na(osmcity), town, osmcity), 
         Newcity = ifelse(is.na(Newcity), municipality, Newcity)) %>%
  relocate(Newcity)


```


###  Some statistics by location with different granularity

We have **`r nrow(mydat)`** files in the directory  *`r SourceDir`* , and **`r nrow(subset(mydatN, Country != ""))`**  with country location. 

```{r}


Tabstat <- mydatN %>%
  group_by(FileType, CountryGPS, Region, Newcity) %>%
  summarize(Nbfiles = n(), .groups = "drop") %>% 
  arrange(desc(Nbfiles)) %>%
  group_split(FileType) %>%
  bind_rows() %>%
  adorn_totals() %>%
  arrange(CountryGPS)

#Note:  .groups = "drop": Stops summarize() from keeping grouping

kable(Tabstat  )

```


## Managing MOV

The idea would be: 

* To move all .mov with *duration > 10s*  to a sub folder
* To rename the .mov *with same name* as the corresponding jpeg 


# Create renaming table

Create, visualize and save table with old and new names. 
## Refine table with specific events


* Identify files corresponding to specific events/places
--> use information on leisure to identify climbing, etc..
--> count pictures per location (fine granularity) and date(dya?)
--> Triggers using Nb of pictures by day/region
* Update renaming table
* Refine to have a minim of files per directory --> Parameter to define at start

# Defining destination directories

The  final naming  structure should be 

 year\_country>month\_Newcity\_year\_month\_day-formername


## Counts of pictures by day, region, city

The 3 main variables for directories and sub-directories are *Year*, *Country* and *Newcity*

```{r Finalcounts}
# If one region, no sub-directory  --> counting that
mydatN <- mydatN %>%
  group_by(Country, Year) %>%
  mutate( NbCity = n_distinct(Newcity)
          )

# If many pictures the same day, then special event sub 
mydatN <- mydatN %>%
  group_by(Date) %>%
  mutate( NbPicturesDay = n_distinct(FileName)) 

# If many pictures in the same Region  then special sub  
mydatN <- mydatN %>%
  group_by(Year,Country,Newcity) %>%
  mutate( NbPicturesCity =  n_distinct(FileName) )

```

##  Final naming decisions

```{r Final naming}
# Defining main rule for Directories and Sub-Directories
mydatN <- mydatN %>%
  mutate(
    MainDir = ifelse(!is.na(Country), Country, "Other"),
    SubDir = Newcity)


# Defining Directories and Sub-Directories
# triggers for SubDirdirectory in final name 
# One case Only for now based on both location and nb of pictures
mydatN <- mydatN %>%
   mutate(SubDir = case_when((NbPicturesDay > NbPicturesDayMax & NbPicturesCity > NbPicturesCityMax)
                             ~ paste0(Month,"_", Newcity),
                             .default =  SubDir)
   )


# Definition of Directories and SubDir

mydatN <-  mydatN  %>%
  mutate(
        Newdir = case_when(
           is.na(SubDir) ~ paste0(Year, "_", MainDir),
           (!is.na(SubDir) &  SubDir !="" ) ~ paste0(Year, "_", MainDir, "/", SubDir, "/"), 
            .default =  paste0(Year, "_", MainDir, "/")
            )
          )
        

```


## Where the `r nrow(mydatN)` pictures will be copied? 

```{r}
# Compiling the number of files that should be in each  new repository
mystat <- mydatN %>%
  group_by(Country, Newdir) %>%
  summarize(NbInNewDir = n(), .groups = "drop") %>%
  arrange(Country, desc(NbInNewDir)) %>%
  group_split(Country) %>%
  adorn_totals()

 kable(mystat)
```

## Adjusting one last time 

**Todo** 

- If  less than  XXX pictures are in a folder then revise  rule!!  But this may be don only **after** a full year has been sorted! 
 


## Creating  origin destination table 


```{r}
# save the file 
library(stringr)
SourceText <- str_replace_all(SourceDir, "[/:]", "_")
SourceText <- gsub("c__", "", SourceText)


toto <- mystat
write.xlsx(,
           file=paste(ToDir,"Log-",SourceText,"-N=", nrow(mydatN), "-(", ymd(Sys.Date()),").xlsx", sep=""))

```


Saving information of old and new names in a csv file **TODO**

We also save this information in a log excel file^[ The file is  `r paste(ToDir,"Log-",SourceText,"-N=", nrow(mydatN), "-(", ymd(Sys.Date()),").xlsx", sep="")` ]



# Remaning (! Danger !) 
Instead of copying, we rename !! 

```{r}
## ACHTUNG !!!!

for(i in 1:nrow(mydatN)) {       # for-loop over rows
   # Create directory if doesn't exist
   ifelse(dir.exists(paste0(ToDir, mydatN[i, "Newdir"])), FALSE,
          dir.create(paste0(ToDir, mydatN[i, "Newdir"]), recursive = TRUE))

  # rename file
   if (file.exists(paste0(SourceDir, mydatN[i, "FileName"]))) {
   file.rename(from = paste0(SourceDir, mydatN[i, "FileName"]),
             to = paste0(ToDir, mydatN[i, "Newdir"],"/", mydatN[i, "NewName"]))
   }
   else {
    warning(paste("File does not exist:", paste0(SourceDir, mydatN[i, "FileName"])))
  }
}

```

## Managing results , veryfing results..
**TODO**
- Store table with full information in general folder --> use later to do maps of where we've been! 
- Store/ identify short.mov somewhere --> Idea: rename to .mov instead of .MOV !!! (think that if we delete a file, well delete the .mov)
- Rename long videos with more information in the name (leisure, name, etc..)
- List (count nb of file in new directory )
- List files not renamed
- Verify counts! 


```{r Copying}
### ACHTUNG !!!! 
# 
# for(i in 1:nrow(mydatN)) {       # for-loop over rows
#    # Create directory if doesn't exist
#    ifelse(dir.exists(paste0(ToDir, mydatN[i, "Newdir"])), FALSE,
#           dir.create(paste0(ToDir, mydatN[i, "Newdir"]), recursive = TRUE))
# 
#   # Copy file
#   file.copy(from = paste0(SourceDir, mydatN[i, "FileName"]),
#             to = paste0(ToDir, mydatN[i, "Newdir"],"/", mydatN[i, "NewName"]))
# 
# }


```


## Checking that everything went fine

We now do a count of the number of pictures in the destination directory

```{r Verif}
MyCopiedfiles <- list.files(path = ToDir,
                    recursive=TRUE,
                    pattern = mypattern, full.names=TRUE)

# Idea is to match the names between Myfiles and MyCopiedfiles

NamesOriginal<- word( gsub("/", " ", Myfiles), -1)     # removing "" from full name and take the last word
NameFinal <- word( gsub("-", " ", MyCopiedfiles), -1)  # <--- Final name has date prefix

NbCommon <- length(intersect(NameFinal, NamesOriginal))

message <- ifelse( NbCommon == nrow(mydatN),
                   paste("Everything went well (", NbCommon , "files copied)"),
                   paste( "**Problem**: there are **", nrow(mydatN) - NbCommon, "** files missing!" )) 

```

### Result of the test: `r message`



# Interactive map with leaflet

```{r}
# text to be displayed on pop-ups

center <- "<div style='text-align:center'>"
name <- paste0(center, "<b>", mydatN$FileName , "</b>", "</div>")
year <- paste0(center,"Date=",  mydatN$Year,"-",mydatN$Month, " </div>")
imgsrc <- paste0( mydat$NewDir, mydat$NewName)
img <- paste0(center, "<img src= '", imgsrc,"' />")

### Not the right order of  images!! 

mydatN %>%
  filter(FileType == "JPEG" & GPSLatitude != 0) %>%
  leaflet() %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addMarkers(~ GPSLongitude, ~ GPSLatitude, popup = paste(name, year, img))  
```

```{r}
library(leaflet)
library(base64enc)
library(dplyr)

# Function to encode an image as base64
encode_img <- function(path) {
  encoded <- base64encode(path)
  paste0("data:image/jpeg;base64,", encoded)
}

# Add encoded images to the data
mydatN <- mydatN %>%
  mutate(
    encoded_img = sapply(file.path(NewDir, NewName), encode_img),
    popup_html = paste0(
      "<div style='text-align:center'>",
      "<b>", FileName, "</b><br>",
      "Date=", Year, "-", Month, "<br>",
      "<img src='", encoded_img, "' width='150'/>",
      "</div>"
    )
  )

# Create Leaflet map
mydatN %>%
  filter(FileType == "JPEG" & GPSLatitude != 0) %>%
  leaflet() %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addMarkers(~GPSLongitude, ~GPSLatitude, popup = ~popup_html)

```


