---
title: "ExifFromPictures"
author: "Christophe Bontemps"
date: "2023-01-29"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE)

# Other options used here 
opts_chunk$set(warning = FALSE, message = FALSE,
               fig.pos = "!H", fig.align = "center",
               autodep = TRUE, tidy = FALSE)
```



```{r libraries}
library(exifr)
library(dplyr)
library(leaflet)
#library("vtable")
library(kableExtra)
library(janitor)


library(openxlsx)
```


```{r exiftoolconfig}
#May be needed to find exiftool (in Windows directory or anywhere in the path)
# From https://exiftool.org/install.html

# library(exiftoolr)
# configure_exiftoolr()
```

## Source and destination
```{r paths}
# Source directory
SourceDir <-  "c:/temp/Photos/Toto"

# Destination 
ToDir <- "c:/Temp/Photos/Titi/"

mypattern  <- "*.jpg|*.JPG|*.PNG|*.MP4|*.MOV"

```

We will transfer the pictures from *`r SourceDir`* to several sub-directories in *`r ToDir`*. 
**NB**: *One idea would be also to do batches of transferts*


##  Extract exif data from pictures
Once exiftool is correctly installed^[(see the information on how to install [https://exiftool.org/install.html](https://exiftool.org/install.html)], one can collect all the information embedded in our pictures. Here we only collect the files with extensions *`r mypattern`*.


```{r exifextraction  }

myjpg <- list.files(path = SourceDir, pattern = mypattern )
files <- list.files(path = SourceDir,
                    recursive=TRUE,
                    pattern = mypattern, full.names=TRUE)

# This is long 
exifinfo <- read_exif(files, quiet = TRUE)
```

We have now the complete information on **`r nrow(exifinfo)`** pictures, among which **`r nrow(subset(exifinfo, FileType =="JPEG"))`** are jpegs. 




```{r selecting variables }

# Using the data 
# https://www.r-bloggers.com/2016/11/extracting-exif-data-from-photos-using-r/
# and https://www.r-bloggers.com/2016/11/create-an-interactive-web-map-with-geotagged-photos/

exifdat <- exifinfo %>%
  select(
    SourceFile, FileName, FileType, FileModifyDate, DateTimeOriginal,
    GPSLongitude, GPSLatitude, GPSAltitude)

```

## Refining the information
We need a bit of cleaning there, in particular, we have date format and, as usual, this is needs to be extracted if one wants to use the year as a sorting variable. Also some coordinates (GPS) are missing and values should be imputed (as zeros here) for *maps* to work (see below).


```{r DateInfo}
library(lubridate)
library(maps)

# Adding date 
mydat <- exifdat %>%
  mutate(DateTimeOriginal = ifelse(is.na(DateTimeOriginal),
                                   FileModifyDate,
                                   DateTimeOriginal)) %>%
  mutate(Date = ymd_hms(DateTimeOriginal)
  ) %>%
  arrange(Date) %>%
 tidyr::separate(Date, c("Year", "rest"), extra = "merge", fill = "left")  %>%
 tidyr::separate(rest, c("Month", "rest2"), extra = "merge", fill = "left") %>%
 tidyr::separate(rest2, c("Day", "Time"),  extra = "merge", fill = "left") %>%
 mutate(
   Date =  paste0(Year,"-",Month, "-", Day),  
   NewName = paste0(Date, "-",  FileName))
```

###  Collecting country information 
We use the package *maps* to recover the information from GPS (latitude, longitude). We will use this information to organize the files into folder names with information on  years and  location. 

```{r CountryInfo}
# Replace missing coordinates by (0,0)
mydat <- mydat %>%
  mutate(GPSLongitude =  ifelse(is.na(GPSLatitude), 0 ,GPSLongitude),
         GPSLatitude =  ifelse(is.na(GPSLatitude), 0 ,GPSLatitude)) 

# Adding Countries from  GPS 
mydat <- mydat %>%
  mutate(Country = maps::map.where(database="world", GPSLongitude, GPSLatitude, na.rm=TRUE),
         Dept = maps::map.where(database="france", GPSLongitude, GPSLatitude, na.rm=TRUE))

# Create CountryShort and Region for detailled information 
mydat <- mydat %>%
  tidyr::extract(Country, c("Countryshort", "Region"),
                 regex = "([[:alnum:]]+):([[:alnum:]]+)",
                 remove = FALSE)%>%
  mutate(Country = ifelse(!is.na(Countryshort) , Countryshort, Country), 
         Region = ifelse(!is.na(Dept) , Dept, Region) )  %>%
 select( !c("Countryshort", "Dept"))

# Identifying specific places/events
### Identifying  P2000
mydat <- mydat %>%
  mutate(Country = ifelse(GPSAltitude >1700, "P2000", Country)) 
 

####  Festa Design
mydat <- mydat %>%
  mutate(Region = ifelse(Date %in% c("2022-11-19", "2022-11-20"), "Design-Festa", Region)) 
  
```


###  Some statistics by location 

We have **`r nrow(mydat)`** files in the directory  *`r SourceDir`* , and **`r nrow(subset(mydat, Country != ""))`**  with country location. 


```{r}

# mydat %>%
#   count(FileType, Country,Region) %%
#    adorn_totals() -> Tabstat


Tabstat <- mydat %>%
  group_by(FileType, Country,Region) %>%
  summarize(Nbfiles = n()) %>%
  adorn_totals()

kable(Tabstat  )

```


# Renaming and Moving pictures to destination directories

```{r Final naming}

# If one region, no sub-directory  --> counting that
mydat <- mydat %>%
  group_by(Country) %>%
  mutate( NbRegion = n_distinct(Region))

# If many pictures the same day, then special event
mydat <- mydat %>%
  group_by(Date) %>%
  mutate( NbPicturesDay = n_distinct(FileName)) 

# Defining Directories and SubDirectories

mydat <- mydat %>%
  mutate(
    Region = ifelse(NbPicturesDay >50, Date, Region), 
    NbRegion = ifelse(NbPicturesDay >50, 2, NbRegion), 
  )

# Definition of Directories 
mydat <-  mydat  %>%
  mutate(
        Newdir = case_when(
          is.na(Country)  ~ paste0(Year, "/",Month),
          !is.na(Country) & is.na(Region) ~ paste0(Year, "/", Country),
          !is.na(Country) & !is.na(Region) & NbRegion == 1  ~ paste0(Year, "/", Country),
          !is.na(Country) & !is.na(Region)  & NbRegion >1 ~ paste0(Year, "/", Country, "/", Region)
          
          )
        )


```

```{r Copying}


for(i in 1:nrow(mydat)) {       # for-loop over rows
   # Create directory if doesn't exist
   ifelse(dir.exists(paste0(ToDir, mydat[i, "Newdir"])), FALSE,
          dir.create(paste0(ToDir, mydat[i, "Newdir"]), recursive = TRUE))

#   # Copy file
#   file.copy(from = paste0(SourceDir, mydat[i, "FileName"]), 
#             to = paste0(ToDir, mydat[i, "Newdir"],"/", mydat[i, "NewName"]))
#   
}


```

# Listing the number of pictures pasted and location

```{r}
#   mUST FIND A WAY
mystat <- mydat %>%
  group_by(Newdir) %>%
  summarize(NbInNewDir = n()) %>%
  adorn_totals()

 kable(mystat)

# save the file 
library(stringr)
SourceText <- str_replace_all(SourceDir, "[/:]", "_")
SourceText <- gsub("c__", "", SourceText)
 
write.xlsx(mystat,
           file=paste(ToDir,"Log-",SourceText,"-N=", nrow(mydat), "-(", ymd(Sys.Date()),").xlsx", sep=""))

```



# Interactive map with leaflet

```{r}
# text to be diplayed on pop-ups

center <- "<div style='text-align:center'>"
name <- paste0(center, "<b>", mydat$FileName , "</b>", "</div>")
year <- paste0(center,"Year=",  mydat$Year, " </div>")
imgsrc <- paste0(SourceDir,"/IMG_0311.JPG")
img <- paste0(center, "<img src= '", imgsrc,"' />")

### Not the right order of  images!! 

mydat %>%
  filter(FileType == "JPEG") %>%
  leaflet() %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addMarkers(~ GPSLongitude, ~ GPSLatitude, popup = paste(name, year, img))  
```


