---
title: "ExifFromPictures"
author: "Christophe Bontemps"
date: "2023-01-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r libraries}
library(exifr)
library(dplyr)
library(leaflet)
```


```{r exiftoolconfig}
#May be needed to find exiftool (in Windows directory or anywhere in the path)
# From https://exiftool.org/install.html

# library(exiftoolr)
# configure_exiftoolr()
```
# Extract exif data from pictures
Once exiftool is correctly installed, one ca collect all the information! 

```{r exifextraction}
DirFiles <-  "c:/temp/Photos/Toto"
mypattern  <- "*.jpg|*.JPG|*.PNG|*.MP4|*.MOV"
myjpg <- list.files(path = DirFiles, pattern = mypattern )

files <- list.files(path = DirFiles,
                    recursive=TRUE,
                    pattern = mypattern, full.names=TRUE)
exifinfo <- read_exif(files, quiet = FALSE)
```



```{r subset }

# Using the data 
# https://www.r-bloggers.com/2016/11/extracting-exif-data-from-photos-using-r/
# and https://www.r-bloggers.com/2016/11/create-an-interactive-web-map-with-geotagged-photos/
mydat <- exifinfo %>%
  select(
    SourceFile, FileName, FileModifyDate, DateTimeOriginal,
    GPSLongitude, GPSLatitude, GPSAltitude)

```

# Collecting country information 

```{r CountryInfo}
library(lubridate)
library(maps)

# Adding date 
mydat <- mydat %>%
  mutate(DateTimeOriginal = ifelse(is.na(DateTimeOriginal),
                                   FileModifyDate, DateTimeOriginal)) %>%
  mutate(Date = ymd_hms(DateTimeOriginal)
  ) %>%
  arrange(Date) %>%
 tidyr::separate(Date, c("Year", "rest"), extra = "merge", fill = "left")  %>%
 tidyr::separate(rest, c("Month", "rest2"), extra = "merge", fill = "left") %>%
 tidyr::separate(rest2, c("Day", "Time"),  extra = "merge", fill = "left") %>%
 mutate( newname = paste0(Year,"-",Month, "-", Day, "-",  FileName))



#  Pictures with GPS 
mydatGPS <- mydat %>%
  filter(!is.na(GPSLatitude)) %>%
  mutate(Country = maps::map.where(database="world", GPSLongitude, GPSLatitude),
         Dept = maps::map.where(database="france", GPSLongitude, GPSLatitude)) %>%
  tidyr::extract(Country, c("Countryshort", "Region"),
                 regex = "([[:alnum:]]+):([[:alnum:]]+)",
                 remove = FALSE)%>%
  mutate(Country = ifelse(!is.na(Countryshort) , Countryshort, Country), 
         Region = ifelse(!is.na(Dept) , Dept, Region) ) %>%
  select( !c("Countryshort", "Dept"))
  
  
```

> we have `r nrow(mydat)` pictures in the directory XXXX, and  `r nrow(mydatGPS)` with country location. 


# Renaming and Moving pictures to directories

```{r GPS moving}
mydatGPS <-  mydatGPS  %>%
  mutate(
         newdir = case_when(is.na(Country)  ~ paste0(Year, "-",Month),
                            !is.na(Country) & is.na(Region) ~ paste0(Year, "-", Country),
                            !is.na(Country) & !is.na(Region) ~ paste0(Year, "-", Region)))





```




# Interactive map with leaflet

```{r}
leaflet(mydat) %>%
addProviderTiles("Esri.WorldImagery") %>%
addMarkers(~ GPSLongitude, ~ GPSLatitude)  
```



```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
